Vertical shear instability in astrophysical disks: vertically global linear calculations with thermal relaxation and applications to the Solar Nebula

### 2015-03-15 setup.tex
added convective overstability discussion
about to edit numerical

For most of this paper we will take $\beta$ as a constant input
parameter, so that we can vary the thermodynamic response of the
disk from isothermal ($\beta\ll 1$) to adiabatic ($\beta \gg 1$) in a
controlled manner. However, for application to protoplanetary disks in 
\S\ref{application} we will model $\beta$ based on physical thermal
timescales, in which case $\beta$ is determined by the equilibrium
disk properties and the perturbation lengthscales of interest.

@@

We consider disk equilibria in which the pressure $P$ and density
$\rho$ are related by   
\begin{align}
  P(r,z) &= 
  \frac{c_{s0}^2(r)}{\Gamma}\rho_0(r)^{1-\Gamma}\rho^\Gamma(r,z),\notag\\
  & \equiv K(r)\rho^{\Gamma}(r,z). \label{eqm_press}
\end{align}
where the polytropic index $\Gamma$ is a  constant parameter  and the
function $c_{s0}(r)$ is prescribed below. Choosing $\Gamma=1$ gives 
locally vertically isothermal disk and $c_{s0}(r)$ is the mid-plane
sound-speed. %For $\Gamma=\gamma$ the disk is adiabatically stratified
% in the vertical 
It shall be convenient to define the function $c_s(r,z)$ 
\begin{align}
  c_s^2\equiv % \frac{dP}{d\rho} =
  c_{s0}^2(r)\left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} =
  \frac{\Gamma P}{\rho}, 
\end{align}
which is related to the isothermal sound-speed $c_\mathrm{iso} =
c_s/\sqrt{\Gamma}$. We take $c_{s0}(r)$ to be another power-law:  
\begin{align}
  c_{s0}^2(r)=c_{s00}^2\left(\frac{r}{r_0}\right)^q, 
\end{align}
where $c_{s00}\equiv c_{s0}(r_0)$  and $q$ is the power-law index, and
define   
\begin{align}
  \epsilon(r) \equiv \frac{c_{s0}(r)}{v_k(r)} \equiv
  \frac{H}{r} 
\end{align}
to parametrize the magnitude of the sound-speed, where
$v_k=r\Omega_k$ is the Keplerian azimuthal velocity and
$H$ is a characteristic scale-height.  Thus, $\epsilon$
is also a characteristic disk aspect-ratio. We are 
interested in thin disks with $\epsilon \ll 1$.%, as relevant for protoplanetary disks \citep{chiang97}.

For discussion purposes it is useful to define the vertical buoyancy
frequency $N_z$ via
\begin{align}
  N_z^2 \equiv - c_p^{-1}g_z \frac{\p S}{\p z} = c_s^2 \left(\frac{\gamma -
      \Gamma}{\gamma}\right)\left(\frac{\p\ln\rho}{\p z}\right)^2,   
\end{align}
where $c_p$ is the heat capacity at constant pressure, $g_z\equiv
\rho^{-1}\p_zP $ and $S\equiv c_p\ln{P^{1/\gamma}/\rho}$ is the
specific entropy. The second equality applies to the disk equilibria
described above. 
We are interested in
neutrally or sub-adiabatically stratified 
disks with $\Gamma\leq \gamma$ so that $N_z^2\geq0$.

@@

\subsection{Vertical shear}\label{vshear_def}
Note that the disk is in general baroclinic with $\Omega =
\Omega(r,z)$. This is explicit from Eq. \ref{eqm_rot} and Eq. \ref{vertiso_eqm}, but is also 
implied by Eq. \ref{vert_equilibrium}---\ref{hori_equilibrium} and 
Eq. \ref{eqm_press},  
\begin{align}
  r\frac{\p\Omega^2}{\p z} = - \rho^{\Gamma-1}\frac{\p\ln\rho}{\p
    z}\frac{dK}{dr} = - \frac{g_z}{\Gamma}\frac{d\ln K}{dr}.\label{vertical_shear}
\end{align}
Evaluating this expression explicitly,
\begin{align}
  r\frac{\p\Omega^2}{\p z} = \frac{sz\Omega_k^2}{\Gamma
    r\left(1+z^2/r^2\right)^{3/2}}. \label{vertical_shear_ex} 
\end{align}
Note that there is a maximum vertical shear rate at $z=r/\sqrt{2}$,
which is expected to limit the growth rate of instabilities associated with vertical shear.

\subsection{Stability condition against axisymmetric adiabatic
  perturbations}\label{solberg}
A rotating flow is stable against axisymmetric adiabatic perturbations
if both Solberg-Hoiland criteria  are satisfied:
\begin{align}
  \kappa^2 - \frac{1}{\rho c_p}\nabla P \cdot \nabla S &> 0,\\
  -\frac{\p P}{\p z}\left(\frac{\p j^2}{\p r}\frac{\p S}{ \p z} -
    \frac{\p j^2}{\p z}\frac{\p S}{\p r} \right)&>0. \label{second_SH} 
\end{align}
\citep{tassoul78}, where $j=r^2\Omega$ is the specific angular
momentum. For rotationally-supported, thin disks, pressure gradients are small 
compared to rotation, so the first criterion is generally
satisfied. Thus, to assess stability, we consider the second
criterion. For our disk models,
\begin{align}
  &\frac{\p S}{\p r} = c_p\left[\frac{s}{\gamma r} +
    \left(\frac{\Gamma}{\gamma}-1\right)\frac{\p \ln{\rho}}{\p
       r}\right],\\
  &\frac{\p S}{\p z} = c_p
  \left(\frac{\Gamma}{\gamma}-1\right)\frac{\p \ln{\rho}}{\p z}. 
\end{align} 
Without loss of generality, we evaluate Eq. \ref{second_SH} in $z>0$
where $\p_zP<0$. Here we approximate $\p_r j^2\simeq r^3\Omega_k^2$ but
use Eq. \ref{vertical_shear} for $\p_zj^2$. Then 
\begin{align}\label{solberg2}
  % \gamma - \Gamma > \frac{s\epsilon^2}{\Gamma}
  % \left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} \left[s
  %   -\left(\gamma-\Gamma\right)\frac{\p\ln{\rho}}{\p\ln{r}} \right]. 
  \gamma - \Gamma > \frac{\epsilon^2}{\Gamma}
  \left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} \left[s^2
    -s\left(\gamma-\Gamma\right)\frac{\p\ln{\rho}}{\p\ln{r}} \right]
\end{align} 
implies stability. For typical model parameters the right hand
side (RHS) is $O(\epsilon^2)$. Hence, instability in the adiabatic limit  
requires $\gamma < \Gamma + O(\epsilon^2)$. Since we consider
$\epsilon\ll1$, this implies the equilibrium disk must have a small
entropy gradient in order to be unstable to axisymmetric adiabatic
perturbations. This is generally not true for passively irradiated, stably
stratified disks \citep{chiang97}, in which the VSI requires rapid
thermal relaxation, as shown below.

@@

The linearized fluid equations in cylindrical co-ordinates are  
\begin{align}
  \ii\sigma\delta\rho &= \frac{\rho}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p}{\p
    z}\left(\rho\delta v_z\right)  + \hat{g}_c\frac{\p\rho}{\p r}\delta v_r,\\
  \ii\sigma\delta v_r &= \frac{1}{\rho}\frac{\p\delta P}{\p r} - 2\Omega
  \delta v_\phi - \hat{g}_c\frac{\delta\rho}{\rho^2}\frac{\p P}{\p r},\\ 
  \ii\sigma\delta v_\phi  &= \frac{\kappa^2}{2\Omega}\delta v_r +
  r\frac{\p\Omega}{\p z}\delta v_z,\\
  \ii \sigma\delta v_z &= \frac{1}{\rho}\frac{\p\delta P}{\p z} -
  \frac{\delta \rho}{\rho^2}\frac{\p P}{\p z},\\
  \ii\sigma\delta P &=\frac{\p P}{\p z}\delta v_z + \gamma P
  \left[\frac{1}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p
      \delta v_z}{\p z}\right]\notag \\
  &+ \frac{1}{t_c}\left(\delta P -  
    \frac{P}{\rho}\delta\rho\right) + \hat{g}_c \frac{\p P}{\p r}\delta v_r,\label{full_energy1}
\end{align}
where the coefficients $\hat{g}_c=1$ are inserted to label terms
associated with the background radial disk structure.

\subsection{Radially-local approximation}  
We consider perturbations with small radial lengthscales 
so that we can ignore the radial variations in the background
structure and radial gradients. We thus evaluate all coefficients in
the above equations at the fiducial radius $r=r_0$. However, the full
vertical dependence of the coefficients are retained. Then we can
assume perturbations have spatial dependence in the form
\begin{align}
  \delta \rho(r,z) \to \delta\rho(z)\exp{\ii k_r r}.
\end{align}
We take $k_r>0$ without loss of generality, and further assume $k_rr\gg1$ so that curvature terms can also be 
ignored.

In order to emphasize the radially-local approximation, we
introduce the local Cartesian co-ordinates $(x,y,z)$ corresponding to
the the local $(r,\phi,z)$ directions, respectively. Then writing
$\delta P /\rho \equiv W$ and $c_s^2\delta\rho/\rho\equiv Q$, the
linearized system of equations are 
\begin{align}
  \ii\sigma \frac{Q}{c_s^2}  &=  \ii k_x \delta v_x + \frac{d\dd
    v_z}{dz} + \frac{\p\ln{\rho}}{\p z}\delta v_z + \hat{g}_c
  \frac{\p\ln{\rho}}{\p r} \delta v_x,\label{lin_mass}\\
  \ii\sigma \delta v_x  &= \ii k_x W - 2\Omega\delta v_y -
  \hat{g}_c\frac{1}{\rho}\frac{\p P}{\p r}\frac{Q}{c_s^2}\label{lin_vx}\\
   \ii \sigma\delta v_y &= r\frac{\p \Omega}{\p z}\delta v_z +
  \frac{\kappa^2}{2\Omega}\delta v_x, \label{lin_vy}\\
   \ii\sigma\delta v_z &= \frac{dW}{dz} +
  \frac{\p\ln{\rho}}{\p z}\left(W-Q\right), \label{lin_vz}\\
  \ii \sigma W &= c_s^2\frac{\p\ln{\rho}}{\p z}\delta v_z +
  c_s^2\frac{\gamma}{\Gamma} \left(\ii k_x\delta v_x + \frac{d\delta
      v_z}{dz}\right) \notag\\
  &+ \frac{1}{t_c}\left(W - \frac{Q}{\Gamma}\right) +
  \hat{g}_c\frac{1}{\rho}\frac{\p P}{\p r}\delta v_x,\label{lin_energy}
  % -&\ii\sigma W + \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right) = -\ii\sigma\frac{\gamma}{\Gamma} Q +
  % c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma}-1\right)\delta v_z.\label{lin_energy}
\end{align}
where we have made the notation change $k_r\to k_x$. 
It is understood that the quantities $\rho,\,c_s^2,\,\Omega,\kappa,\, 
\p_z\Omega,\, \p_r\ln{\rho} $ and $ \p_r P$ are evaluated at $r=r_0$
using their expressions in the global disk given in \S\ref{setup}. Thus, they are functions of $z$ only. 
Eq. \ref{lin_mass}---\ref{lin_energy} will form the basis of our numerical
study, but for analytical discussion
we will make further simplifications.

% The radially-local approximation above is similar in spirit to the
% shearing box framework \citep{goldreich65} or the generalization
% thereof \citep{mcnally14}. The main difference being the terms
% proportional to $\hat{g}_c$ are neglected in the shearing box.

>>>>

For most of this paper we will take $\beta$ as a constant input
parameter, so that we can vary the thermodynamic response of the
disk from isothermal ($\beta\ll 1$) to adiabatic ($\beta \gg 1$) in a
controlled manner. However, for application to protoplanetary disks in 
\S\ref{application} we will model $\beta$ based on physical thermal
timescales, in which case $\beta$ is non-constant and is determined by
the equilibrium disk properties and the perturbation lengthscales of
interest.

@@

We consider disk equilibria in which the pressure $P$ and density
$\rho$ are related by   
\begin{align}
  P(r,z) &= 
  \frac{c_{s0}^2(r)}{\Gamma}\rho_0(r)^{1-\Gamma}\rho^\Gamma(r,z),\notag\\
  & \equiv K(r)\rho^{\Gamma}(r,z). \label{eqm_press}
\end{align}
where the polytropic index $\Gamma$ is a  constant parameter  and the
function $c_{s0}(r)$ is prescribed below. Choosing $\Gamma=1$ gives  a 
locally vertically isothermal disk and $c_{s0}(r)$ is the mid-plane
sound-speed. 
It shall be convenient to define the function $c_s(r,z)$ 
\begin{align}
  c_s^2\equiv % \frac{dP}{d\rho} =
  c_{s0}^2(r)\left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} =
  \frac{\Gamma P}{\rho}, 
\end{align}
which is related to the isothermal sound-speed $c_\mathrm{iso} =
c_s/\sqrt{\Gamma}$. We take $c_{s0}(r)$ to be another power-law:  
\begin{align}
  c_{s0}^2(r)=c_{s00}^2\left(\frac{r}{r_0}\right)^q, 
\end{align}
where $c_{s00}\equiv c_{s0}(r_0)$  and $q$ is the power-law index, and
define   
\begin{align}
  \epsilon(r) \equiv \frac{c_{s0}(r)}{v_k(r)} \equiv
  \frac{H}{r} 
\end{align}
to parametrize the magnitude of the sound-speed, where
$v_k=r\Omega_k$ is the Keplerian azimuthal velocity and
$H$ is a characteristic scale-height.  Thus, $\epsilon$
is also a characteristic disk aspect-ratio. We are 
interested in thin disks with $\epsilon \ll 1$.%,%  as relevant for
% protoplanetary disks \citep{chiang97}.

For discussion purposes it is useful to define the vertical buoyancy
frequency $N_z$ via
\begin{align}
  N_z^2 \equiv - c_p^{-1}g_z \frac{\p S}{\p z} = c_s^2 \left(\frac{\gamma -
      \Gamma}{\gamma}\right)\left(\frac{\p\ln\rho}{\p z}\right)^2,   \label{nzsq_def}
\end{align}
where $c_p$ is the heat capacity at constant pressure, $g_z\equiv
\rho^{-1}\p_zP $ and $S\equiv c_p\ln{P^{1/\gamma}/\rho}$ is the
specific entropy. The second equality applies to the disk equilibria
described above. 
We are interested in
neutrally or sub-adiabatically stratified 
disks with $\Gamma\leq \gamma$ so that $N_z^2\geq0$.

@@

\subsection{Vertical shear}\label{vshear_def}
Note that the disk is in general baroclinic with $\Omega =
\Omega(r,z)$. This is explicit from Eq. \ref{eqm_rot} and Eq. \ref{vertiso_eqm}, but is also 
implied by Eq. \ref{vert_equilibrium}---\ref{hori_equilibrium} and 
Eq. \ref{eqm_press},  
\begin{align}
  r\frac{\p\Omega^2}{\p z} = - \rho^{\Gamma-1}\frac{\p\ln\rho}{\p
    z}\frac{dK}{dr} = - \frac{g_z}{\Gamma}\frac{d\ln K}{dr}.\label{vertical_shear}
\end{align}
For the above disk models,
\begin{align}
  r\frac{\p\Omega^2}{\p z} = \frac{sz\Omega_k^2}{\Gamma
    r\left(1+z^2/r^2\right)^{3/2}}. \label{vertical_shear_ex} 
\end{align}
Note that there is a maximum vertical shear rate at $z=r/\sqrt{2}$,
which is expected to limit the growth rate of instabilities associated with vertical shear.

\subsection{Stability condition against axisymmetric adiabatic
  perturbations}\label{solberg}
A rotating flow is stable against axisymmetric adiabatic perturbations
if both Solberg-Hoiland criteria  are satisfied:
\begin{align}
  \kappa^2 - \frac{1}{\rho c_p}\nabla P \cdot \nabla S &> 0,\\
  -\frac{\p P}{\p z}\left(\frac{\p j^2}{\p r}\frac{\p S}{ \p z} -
    \frac{\p j^2}{\p z}\frac{\p S}{\p r} \right)&>0. \label{second_SH} 
\end{align}
\citep{tassoul78}, where $j=r^2\Omega$ is the specific angular
momentum. For rotationally-supported, thin disks %  pressure gradients are small 
% compared to rotation, so 
the first criterion is generally satisfied. 
Thus, to assess stability, we consider the second
criterion. For our disk models,
\begin{align}
  &\frac{\p S}{\p r} = c_p\left[\frac{s}{\gamma r} +
    \left(\frac{\Gamma}{\gamma}-1\right)\frac{\p \ln{\rho}}{\p
       r}\right],\\
  &\frac{\p S}{\p z} = c_p
  \left(\frac{\Gamma}{\gamma}-1\right)\frac{\p \ln{\rho}}{\p z}. 
\end{align} 
Without loss of generality, we evaluate Eq. \ref{second_SH} in $z>0$
where $\p_zP<0$. Here we approximate $\p_r j^2\simeq r^3\Omega_k^2$ but
use Eq. \ref{vertical_shear} for $\p_zj^2$. Then 
\begin{align}\label{solberg2}
  % \gamma - \Gamma > \frac{s\epsilon^2}{\Gamma}
  % \left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} \left[s
  %   -\left(\gamma-\Gamma\right)\frac{\p\ln{\rho}}{\p\ln{r}} \right]. 
  \gamma - \Gamma > \frac{\epsilon^2}{\Gamma}
  \left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} \left[s^2
    -s\left(\gamma-\Gamma\right)\frac{\p\ln{\rho}}{\p\ln{r}} \right]
\end{align} 
implies stability. For typical model parameters the right hand
side (RHS) is $O(\epsilon^2)$. Hence, instability in the adiabatic limit  
requires $\gamma < \Gamma + O(\epsilon^2)$. Since we consider
$\epsilon\ll1$, this implies the equilibrium disk must have a small
entropy gradient in order to be unstable to axisymmetric adiabatic
perturbations. This is generally not true for passively irradiated, stably
stratified disks \citep{chiang97}, in which the VSI requires rapid
thermal relaxation, as shown below.

@@

The linearized fluid equations in cylindrical co-ordinates are  
\begin{align}
  \ii\sigma\delta\rho &= \frac{\rho}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p}{\p
    z}\left(\rho\delta v_z\right)  + \hat{g}_c\frac{\p\rho}{\p r}\delta v_r,\label{full_mass1}\\
  \ii\sigma\delta v_r &= \frac{1}{\rho}\frac{\p\delta P}{\p r} - 2\Omega
  \delta v_\phi - \hat{g}_c\frac{\delta\rho}{\rho^2}\frac{\p P}{\p r},\\ 
  \ii\sigma\delta v_\phi  &= \frac{\kappa^2}{2\Omega}\delta v_r +
  r\frac{\p\Omega}{\p z}\delta v_z,\\
  \ii \sigma\delta v_z &= \frac{1}{\rho}\frac{\p\delta P}{\p z} -
  \frac{\delta \rho}{\rho^2}\frac{\p P}{\p z},\\
  \ii\sigma\delta P &=\frac{\p P}{\p z}\delta v_z + \gamma P
  \left[\frac{1}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p
      \delta v_z}{\p z}\right]\notag \\
  &+ \frac{1}{t_c}\left(\delta P -  
    \frac{P}{\rho}\delta\rho\right) + \hat{g}_c \frac{\p P}{\p r}\delta v_r,\label{full_energy1}
\end{align}
where the coefficients $\hat{g}_c=1$ are inserted to label terms
associated with the background radial disk structure.

\subsection{Radially-local approximation}  
We consider perturbations with small radial lengthscales 
so that we can ignore the radial variations in the background
structure and radial gradients. We thus evaluate all coefficients in
Eq. \ref{full_mass1}---\ref{full_energy1} at the fiducial radius
$r=r_0$. However, the full vertical dependence of the coefficients are
retained. Then we can assume perturbations have spatial dependence in
the form 
\begin{align}
  \delta \rho(r,z) \to \delta\rho(z)\exp{\ii k_r r}.
\end{align}
We take $k_r>0$ without loss of generality, and further assume $k_rr\gg1$ so that curvature terms can also be 
ignored.

In order to emphasize the radially local approximation, we
introduce the local Cartesian co-ordinates $(x,y,z)$ corresponding to
the the local $(r,\phi,z)$ directions, respectively. Then writing
$\delta P /\rho \equiv W$ and $c_s^2\delta\rho/\rho\equiv Q$, the
linearized system of equations are 
\begin{align}
  \ii\sigma \frac{Q}{c_s^2}  &=  \ii k_x \delta v_x + \frac{d\dd
    v_z}{dz} + \frac{\p\ln{\rho}}{\p z}\delta v_z + \hat{g}_c
  \frac{\p\ln{\rho}}{\p r} \delta v_x,\label{lin_mass}\\
  \ii\sigma \delta v_x  &= \ii k_x W - 2\Omega\delta v_y -
  \hat{g}_c\frac{1}{\rho}\frac{\p P}{\p r}\frac{Q}{c_s^2}\label{lin_vx}\\
   \ii \sigma\delta v_y &= r\frac{\p \Omega}{\p z}\delta v_z +
  \frac{\kappa^2}{2\Omega}\delta v_x, \label{lin_vy}\\
   \ii\sigma\delta v_z &= \frac{dW}{dz} +
  \frac{\p\ln{\rho}}{\p z}\left(W-Q\right), \label{lin_vz}\\
  \ii \sigma W &= c_s^2\frac{\p\ln{\rho}}{\p z}\delta v_z +
  c_s^2\frac{\gamma}{\Gamma} \left(\ii k_x\delta v_x + \frac{d\delta
      v_z}{dz}\right) \notag\\
  &+ \frac{1}{t_c}\left(W - \frac{Q}{\Gamma}\right) +
  \hat{g}_c\frac{1}{\rho}\frac{\p P}{\p r}\delta v_x,\label{lin_energy}
  % -&\ii\sigma W + \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right) = -\ii\sigma\frac{\gamma}{\Gamma} Q +
  % c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma}-1\right)\delta v_z.\label{lin_energy}
\end{align}
where we have made the notation change $k_r\to k_x$. 
It is understood that the quantities $\rho,\,c_s^2,\,\Omega,\kappa,\, 
\p_z\Omega,\, \p_r\ln{\rho} $ and $ \p_r P$ are evaluated at $r=r_0$
using their expressions in the global disk given in \S\ref{setup}. Thus, they are functions of $z$ only. 
Eq. \ref{lin_mass}---\ref{lin_energy} will form the basis of our numerical
study, but for analytical discussion
we will make further simplifications.

### 2015-03-15 summary.tex
about to edit numerical

\subsubsection{Parameter study and convective overstability}
We have focused on parameter regimes relevant to the VSI,
namely $\khat\gtrsim 1$ and $\beta\ll 1$. This excludes the
`convective overstability' recently discussed by \cite{klahr14} and
\cite{lyra14} in vertically unstratified disk models.

@@

Evaluating Eq. \ref{max_conv_gen} at the midplane of our disk models,
we find  
\begin{align}
  \nu_\mathrm{conv,max} =
  -\frac{\epsilon^2\Omega_k}{4\gamma\Gamma}\left(p+q\right)
  \left[\left(\gamma-1\right)p-q\right]. \label{max_conv}
\end{align}
% We see that convective overstability is generally slower 
% than the VSI because $\nu_\mathrm{conv,max} = O(\epsilon^2\Omega_k)$,
% whereas the maximum VSI growth rate is $O(\epsilon\Omega_k)$. 
In fact, for the MMSN considered in this work (\S\ref{application} and Appendix
\ref{mmsn}), convective overstability does not apply because
$N_r^2>0$.

However, for more general disk structures convective overstability is
possible. The models adopted for our numerical study in
\S\ref{numerical} do satisfy $N_r^2<0$, but convective overstability
requires considering perturbations with long radial lengthscales,
$\khat\ll 1$, for which radially global disk models would be more 
appropriate.  
% We will explore convective overstability more fully in a
% follow-up study.

\subsubsection{Improved thermal evolution} 
We assumed non-adiabatic effects
can be represented by a thermal relaxation or cooling timescale $t_c$. For 
application to protoplanetary disks we connected $t_c$ to the disk
properties. One one can improve the realism further by generalizing the 
opacity model adopted here to be density (and hence height) dependent. 
%here to be density-dependent,  i.e. $\kappa_d =
%% \kappa_d(\rho, T)$. 
However, going beyond thermal relaxation requires a
proper treatment for radiative diffusion. The linear problem then
involves a Laplacian operator, which require additional boundary
conditions. We will consider these improvements in a
follow-up study. %additional heat source

>>>>

\subsubsection{Convective overstability}
We have focused on parameter regimes relevant to the VSI,
namely $\khat\gtrsim 1$ and $\beta\ll 1$. This excludes the
`convective overstability' recently discussed by \cite{klahr14} and
\cite{lyra14} in vertically unstratified disk models.

@@

Evaluating Eq. \ref{max_conv_gen} at the midplane of our disk models,
we find  
\begin{align}
  \nu_\mathrm{conv,max} =
  -\frac{\epsilon^2\Omega_k}{4\gamma\Gamma}\left(p+q\right)
  \left[\left(\gamma-1\right)p-q\right]. \label{max_conv}
\end{align}
In fact, for the MMSN considered in this work, convective
overstability does not apply because 
$N_r^2>0$.

For more general disk structures convective overstability is
possible. The models adopted for our numerical study in
\S\ref{numerical} do satisfy $N_r^2<0$, but convective overstability
requires considering perturbations with long radial lengthscales,
$\khat\ll 1$, for which radially global disk models would be more 
appropriate.

\subsubsection{Improved thermal evolution} 
We assumed non-adiabatic effects
can be represented by a thermal relaxation or cooling timescale $t_c$. For 
application to protoplanetary disks we connected $t_c$ to the disk
properties. One can improve the realism further by generalizing the 
opacity model adopted here to be density (and hence height) dependent. 
%here to be density-dependent,  i.e. $\kappa_d =
%% \kappa_d(\rho, T)$. 
However, going beyond thermal relaxation requires a
proper treatment of radiative diffusion. The linear problem then
involves a Laplacian operator, which require additional boundary
conditions. We will consider these improvements in a
follow-up study. %additional heat source
